---
title: "Shiny app - prep"
author: "Amelia Ritger"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

```{r, message=FALSE, warning=FALSE}
#Read in data
reef <- read_csv("MBONReef_Histogram.csv")

#Tidy up the data
reef_tidy <- reef %>%
  clean_names() %>% #standardize names
  pivot_longer("annelida_cirriformia_luxuriosa":"substrate_amphipod_tube_complex") %>%  #make long form
  separate(name, into="phylum", sep="_", remove=FALSE) %>% #Add column for phylum name
  mutate(vectorized_name=str_split(name, pattern="_")) %>% #In case this is useful...
  filter(!phylum=="no") %>% #remove values related to data collection issue
  filter(!phylum=="substrate") #remove substrate values

#Because it's faster to do it outside tidyverse
reef_tidy$binary <- ifelse(reef_tidy$value>0, 1, 0) #Add presence/absence column
reef_tidy$species <- gsub("^[^_]*_","",reef_tidy$name, perl=TRUE) #Add column for species name

#Extract only species present
reef_tidy <- reef_tidy %>%
  filter(binary > "0")
```

### Let's try plotting something

#### The counts of each phylum (or species) occurring at each location:
```{r}
ggplot(reef_tidy, aes(x=phylum)) +
  geom_bar(aes(phylum)) +
  facet_wrap(~location)
```

I will allow people to pick their phylum (or species) of interest and pick their location of interest

### Let's try plotting something else

#### The counts of co-occuring phyla (or species) with a focal phylum (or species)
```{r}
reef_select <- reef_tidy %>%
  mutate(focal_phylum="annelida") %>% #pick a focal phylum (BASED ON INPUT)
  mutate(presence = ifelse(phylum==focal_phylum, filename, "FALSE")) %>% #create a column that we can subset all rows in a plot based on the presence of focal phylum in the plot at least once
  filter(filename %in% presence) %>% #if focal phylum is present, keep all observations of that plot ("filename")
  #filter(!phylum=="annelida") %>% #remove the focal phylum, because we don't need to see co-occurrence of focal phylum with focal phylum (BASED ON INPUT)
  filter(phylum==c("porifera", "mollusca")) #select only the coocurring species you want to look at
  
ggplot(reef_select, aes(x=phylum)) +
  geom_bar(aes(phylum)) +
  coord_flip()
```
I will allow people to pick their focal phylum and co-occuring phylum
